Seshadri constants of curve configurations on surfaces

### 2024-03-20 sc-curve-config.tex
Generalised version-2
Gen_Version (12:41)

We first introduce some notation.

	%\begin{definition}
		Let $\mathcal{C}$ be a transversal arrangement of curves on a smooth projective surface $X$.

@@

	\begin{question}\label{question}
		Let $\mathcal{C} =\{C_1,C_2, \ldots ,C_d\}$ be a transversal arrangement of curves on a smooth projective surface $X$. Let $L$ be an ample line bundle on $X$. Is it true that
		\begin{equation*}\varepsilon(X, L,{\rm Sing}(\mathcal{C})) =  \inf_{C_i}{\left{\frac{L\cdot C_i}{b_i}\}\right}?
		\end{equation*}
	\end{question}

By \cite[Example 3.5]{JP}, the above question has a negative answer for the Hesse arrangement of conics 
in $\mathbb{P}^2$. However it is open for line arrangements in $\mathbb{P}^2$. 
It is interesting to study the situations in which it is true on arbitrary surfaces.

	For star arrangements, the base constant is exactly equal to $C_1^2(d-1)$. We answer this question positively in Corollary \ref{prop:star} for some arrangement.

The following is our main theorem. It gives the value of Seshadri constant for an arbitrary transversal configuration 
$\mathcal{C} = \{C_1, \dots, C_d\}$
under a nefness condition.

@@

@@

 \begin{remark} 
Recall that the arrangements we study in this paper are all connected. Since each curve in $\mathcal{C}$ is assumed to be linearly equivalent to a fixed divisor $A$, 
Assumption \ref{star1} implies in particular that $C_i\cdot C_j > 0$ for all $1\le i,j \le d$.  
 \end{remark}
	\begin{definition}
		Let $\mathcal{C}$ be a transversal arrangement of curves on a smooth projective surface $X$
		satisfying Assumption \ref{star1}. Then we define the \emph{base constant} of $\mathcal{C}$ as 
		$${\rm bs}(\mathcal{C}) := {\rm max} \{t \, | \, t = \# \, C_{i} \cap {\rm Sing}(\mathcal{C}), \, C_{i} \in \mathcal{C}\},$$
		i.e., ${\rm bs}(\mathcal{C})$ is equal to the maximum number of singular points  of $\mathcal{C}$ that are contained in a single curve $C_{i} \in \mathcal{C}$.
	\end{definition}

	\begin{question}\label{question}
		Let $\mathcal{C} =\{C_1,C_2, \ldots ,C_d\}$ be a transversal arrangement of curves on a smooth projective surface $X$
		satisfying Assumption \ref{star1}. Let $L$ be an ample line bundle on $X$. Is it true that
		\begin{equation*}\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \frac{L \cdot C_1}{{\rm bs}(\mathcal{C})}?
		\end{equation*}
	\end{question}

By \cite[Example 3.5]{JP}, the above question has a negative answer for the Hesse arrangement of conics 
in $\mathbb{P}^2$. However it is open for line arrangements in $\mathbb{P}^2$. 
It is interesting to study the situations in which it is true on arbitrary surfaces.

@@

	For star arrangements, the base constant is exactly equal to $C_1^2(d-1)$. We answer this question positively in Corollary \ref{prop:star} for some arrangement.

We now prove some corollaries for star arrangements where all the curves are linearly equivalent to a fixed divisor. This answers Question \ref{QuestM} in the affirmative.

@@

\begin{remark}
Several authors have studied multi-point Seshadri constants on surfaces; for a sample of results, see \cite{B,HM, HR, PR, RR, SS}. 
Many of these results consider general points. 
Theorem \ref{prop:star} gives a computation of Seshadri constants at \textit{special} points. 
\end{remark}

The proof of Theorem \ref{prop:star} shows that for an arbitrary arrangement, not necessarily a star arrangement, we have the following lower and upper bounds for the Seshadri constants.

\begin{corollary}\label{cor-main}
Let $X$ be a nonsingular projective surface and let $\mathcal{C} = \{C_{1}, \ldots, C_{d}\}$ be a configuration of curves on $X$ satisfying Assumption \ref{star1}.
%such that each $ C_i$ is linearly equivalent to a fixed divisor $A$ on $X$. 
 Let $L$ be an ample line bundle on $X$
		such that  $\frac{dC_1^2}{(d-1)(L\cdot C_1)}L-C_1$ is nef.   Then
$$\frac{L\cdot C_1}{C_1^2(d-1)} \le \varepsilon(X, L,{\rm Sing}(\mathcal{C})) \le \frac{L \cdot C_1}{{\rm bs} (\mathcal{C})}  .$$
\end{corollary} 
\begin{proof}
The inequality $\frac{L\cdot C_1}{C_1^2(d-1)} \le \varepsilon(X, L,{\rm Sing}(\mathcal{C}))$ follows from the proof of Theorem \ref{prop:star}. Note that for any reduced and irreducible curve $D$ which is not in $\mathcal{C}$, we showed that the Seshadri ratio is at least $\frac{L\cdot C_1}{C_1^2(d-1)}$ and this part does not use the fact that $\mathcal{C}$ is a star arrangement.

@@

It is easy to see that $\frac{24}{11}L - C_1$ is nef. 
Then Corollary \ref{cor-main} implies
$$\frac{1}{22} <  \varepsilon(\mathbb{P}^2, L,{\rm Sing}(\mathcal{C})) = \frac{1}{5}<  \frac{1}{4}.$$

This shows that the inequalities in Corollary \ref{cor-main} are strict, in general. In particular,  Theorem \ref{prop:star} does not hold in this case. Note that $\mathcal{C}$ is not a star arrangement.

>>>>

We first introduce some notation. 
	%\begin{definition}
		Let $\mathcal{C}$ be a transversal arrangement of curves on a smooth projective surface $X$.

@@

	\begin{question}\label{question}
		Let $\mathcal{C} =\{C_1,C_2, \ldots ,C_d\}$ be a transversal arrangement of curves on a smooth projective surface $X$. Let $L$ be an ample line bundle on $X$. Is it true that
		\begin{equation*}
  \varepsilon(X, L,{\rm Sing}(\mathcal{C})) =  \min_{i}\left\{\frac{L\cdot C_i}{b_i}\right\}?
		\end{equation*}
	\end{question}

By \cite[Example 3.5]{JP}, the above question has a negative answer for the Hesse arrangement of conics 
in $\mathbb{P}^2$. However it is open for line arrangements in $\mathbb{P}^2$. 
It is interesting to study the situations in which it is true on arbitrary surfaces. Our main theorem answers this question positively under an additional nefness condition.

%The following is our main theorem. It gives the value of Seshadri  constant for an arbitrary transversal configuration  $\mathcal{C} = \{C_1, \dots, C_d\}$ under a nefness condition.

@@

\begin{remark}
Several authors have studied multi-point Seshadri constants on surfaces; for a sample of results, see \cite{B,HM, HR, PR, RR, SS}. 
Many of these results consider general points. 
Theorem \ref{main-theorem} gives a computation of Seshadri constants at \textit{special} points. 
\end{remark}

@@

 \begin{remark} 
Recall that the arrangements we study in this paper are all connected. Since each curve in $\mathcal{C}$ is assumed to be linearly equivalent to a fixed divisor $A$, 
Assumption \ref{star1} implies in particular that $C_i\cdot C_j > 0$ for all $1\le i,j \le d$.  
 \end{remark}

@@

	For star arrangements, the base constant is exactly equal to $C_1^2(d-1)$. We answer this question positively in the following  corollaries for some arrangements.

%We now prove some corollaries for star arrangements where all the curves are linearly equivalent to a fixed divisor. This answers Question \ref{QuestM} in the affirmative.

@@

The proof of Theorem \ref{main-theorem} shows that for arrangements
satisfying Assumption \ref{star1}, we have the following lower and upper bounds for the Seshadri constants.

\begin{corollary}\label{cor-main}
Let $X$ be a nonsingular projective surface and let $\mathcal{C} = \{C_{1}, \ldots, C_{d}\}$ be a configuration of curves on $X$ satisfying Assumption \ref{star1}.
%such that each $ C_i$ is linearly equivalent to a fixed divisor $A$ on $X$. 
 Let $L$ be an ample line bundle on $X$
		such that  $\frac{dC_1^2}{(d-1)(L\cdot C_1)}L-C_1$ is nef.   Then
$$\frac{L\cdot C_1}{C_1^2(d-1)} \le \varepsilon(X, L,{\rm Sing}(\mathcal{C})) \le \frac{L \cdot C_1}{{\rm bs} (\mathcal{C})}  .$$
\end{corollary} 
\begin{proof}

Note that the hypotheses of Theorem \ref{main-theorem} hold here. 
 For any reduced and irreducible curve $D$ which is not in $\mathcal{C}$, we showed in the proof of Theorem \ref{main-theorem}
that the Seshadri ratio is at least $\frac{L\cdot C_1}{C_1^2(d-1)}$. 
So we obtain $\frac{L\cdot C_1}{C_1^2(d-1)} \le \varepsilon(X, L,{\rm Sing}(\mathcal{C}))$.

@@

We note that the nefness assumption of Theorem \ref{main-theorem} does not hold here. Indeed, $b_i = 8$ for each $1 \le i \le 12$. So the divisor 
$\left(\frac{12(b_1+\dots + b_{12})}{11*11*24}\right)L - C_1$ is not nef because 
$b_1+\dots + b_{12} < 4*11^2$.

However, Corollary \ref{cor-main} does apply in this case. 
It is easy to see that $\frac{24}{11}L - C_1$ is nef. 
Then Corollary \ref{cor-main} implies
$$\frac{1}{22} <  \varepsilon(\mathbb{P}^2, L,{\rm Sing}(\mathcal{C})) = \frac{1}{5}<  \frac{1}{4}.$$

This shows that the inequalities in Corollary \ref{cor-main} are strict, in general. In particular,  Corollary \ref{prop:star} does not hold in this case. Note that $\mathcal{C}$ is not a star arrangement.
